
---
# This RMD file loads the clustered data and finds markers - differentially expressed genes
# Author : Vijay Nagarajan, NEI/NIH
title: "Differential Expression"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE}
options(knitr.duplicate.label = "allow")
knitr::opts_chunk$set()
#knitr::purl(input="/data/../TotalSeq/Tools/differential.Rmd", output="/data/../TotalSeq/Tools/differential.R")
#knitr::purl(input="/data/../TotalSeq/Tools/differential.Rmd", output="/data/../TotalSeq/Tools/differential.R")
```

## 1. Introduction
This document takes the QCd, normalized, scaled, batch corrected, clustered data and performs differential expression analysis.
```{r packages, message=FALSE, warning=FALSE}
#Install and load required R packages. Install the packages if you do not already have them installed
#BiocManager::install('multtest')
#install.packages('metap')
library(Seurat)
library(kableExtra)
library(ggplot2)
library(plotly)
library(tidyverse)
library(rmarkdown)
library(cowplot)

```

## 2. Load QC'd, normalized, scaled, batch corrected data and clustered data
```{r dataload, message=FALSE, warning=FALSE}
# Load QCd, normalized data
load('/data/../TotalSeq/Results/clustering/SeuratObjectAfterClustering.RData')
# Input path
datadirectory <- "/data/../TotalSeq/Results/clustering/"
# Output path
outputpath="/data/../TotalSeq/Results/markers/"
# Set working directory
#setwd(outputpath)
```

## 3. Differential expression (HC vs BCR-UV)
```{r differentialhcvsbcruv, message=FALSE, warning=FALSE}
########################### Marker genes between groups
# Current identity levels
levels(allsamples)
head(Idents(allsamples))
# Change identity to HC vs BCR-UV
Idents(allsamples) <- "type"
head(Idents(allsamples))
levels(allsamples)
########################### tables
# Proportion and counts of cells in clusters and groups
prop.table(table(Idents(allsamples)))
table(Idents(allsamples))
# Find differentially expressed genes for UV vs BCR-UV
markershcbcruv = FindMarkers(allsamples, ident.1="BCR-UV", ident.2="HC")
head(markershcbcruv)
# Set marker rownames
markershcbcruv$gene=rownames(markershcbcruv)
# Save the list of differentially expressed genes as a tab delimited text file
write.table(markershcbcruv, file=paste(outputpath,"markershcbcruv.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")
VlnPlot(allsamples,features=c("HLA-B","TMSB4X"), pt.size = 0)
FeaturePlot(allsamples, features="HLA-B")
table(allsamples$type)

# Get all the expression count matrix
#head(GetAssayData(WT_TG_Filt_Scaled, slot="counts"))
# Get expression values for a single gene from all cells
head(FetchData(allsamples, vars="HLA-B"))
```

## 4. Differential expression between clusters
```{r differentialbetweenclusters, message=FALSE, warning=FALSE}
########################### tables
# Proportion and counts of cells in clusters and groups
# Change identity to 0.25 cluster resolution
Idents(allsamples) <- "RNA_snn_res.0.25"
prop.table(table(Idents(allsamples)))
table(Idents(allsamples), allsamples$type)

########################### Marker genes between clusters
# Current identity levels
levels(allsamples)
head(Idents(allsamples))

# Find differentially expressed genes for cluster 0 vs 1
markerscluster = FindMarkers(allsamples, ident.1=0, ident.2=1)
head(markerscluster)
write.table(markerscluster, file=paste(outputpath,"markerscluster.txt",sep=""), quote=F, row.names = FALSE, sep = "\t")
VlnPlot(allsamples,features=c("TNFRSF4","TNFRSF25"), pt.size = 0)
FeaturePlot(allsamples, features="TNFRSF25")
table(allsamples$orig.ident)

# Get all the expression count matrix
#head(GetAssayData(WT_TG_Filt_Scaled, slot="counts"))
# Get expression values for a single gene from all cells
head(FetchData(allsamples, vars="TNFRSF25"))
```

## 5. Differential expression between every clusters
```{r differentialbetweeneveryclusters, message=FALSE, warning=FALSE}
########################### Find all markers
markersall=FindAllMarkers(allsamples)
head(markersall)
#View(markersall)
# Save all markers as tab delimited file
write.table(markersall, file=paste(outputpath,"markersall.txt",sep=""),quote=F, row.names = FALSE, sep = ",")
# Plot selected markers
VlnPlot(allsamples,features=c("CYP1B1","IL1B"), pt.size = 0)
FeaturePlot(allsamples, features="IL1B")

# Extract the top marker for each cluster
head(markersall)
markersall %>%
  group_by(cluster) %>%
    slice(1) %>%
      pull(gene) -> best.gene.per.cluster

best.gene.per.cluster

# Store plot as ggplot list
vlnplot1<-VlnPlot(allsamples,features=best.gene.per.cluster, pt.size = 0, combine = FALSE)
# vlnplot1<-VlnPlot(allsamples,features=best.gene.per.cluster, pt.size = 0, same.y.lims = TRUE, combine = FALSE)

# Set properties for each plot in the list
p1 <- list()
for (i in seq_along(vlnplot1)){
    #Change x and y tick label font size.
    p1[[i]] = vlnplot1[[i]] +
      theme(legend.position = "none",
            axis.text.x = element_text(size = 8),
            axis.text.y = element_text(size = 8),
            axis.title.y = element_text(size = 6),
            axis.title.x = element_blank(),
            title = element_text(size=8))
}
# Plot all the plots in the list
plot_grid(plotlist = p1, ncol = 3)
# Save plot to file
vlnplot1=plot_grid(plotlist = p1, ncol = 3)
ggsave(paste(outputpath,"vlnplot-best-per-cluster.pdf",sep=""), plot = vlnplot1, width = 20, height = 15, units = "cm")

# Plot features in umap
fplot1<-FeaturePlot(allsamples,features=best.gene.per.cluster, pt.size = 0, combine = FALSE)
p2 <- list()
for (i in seq_along(fplot1)){
    #Change x and y label font size.
    p2[[i]] = fplot1[[i]] +
      theme(legend.position = "none",
            axis.text.x = element_text(size = 8),
            axis.text.y = element_text(size = 8),
            axis.title.y = element_blank(),
            axis.title.x = element_blank(),
            title = element_text(size=8))
}
plot_grid(plotlist = p2, ncol = 4)
fplot1=plot_grid(plotlist = p2, ncol = 4)
ggsave(paste(outputpath,"fplot-best-per-cluster.pdf",sep=""), plot = fplot1, width = 20, height = 15, units = "cm")


```

## 6. Differential expression groups within a specific cluster

```{r differentialgroupsinacluster, message=FALSE, warning=FALSE}
########################### tables
# Proportion and counts of cells in clusters and groups
# Change identity to 0.25 cluster resolution
Idents(allsamples) <- "RNA_snn_res.0.25"
prop.table(table(Idents(allsamples)))
table(Idents(allsamples))

# Current identity levels
levels(allsamples)
head(Idents(allsamples))

# Take all cells in cluster 1, and find markers that separate 'HC' and 'BCR-UV' that are in the 'orig.ident' column. subset.ident would the default identity of clusters
markersingroup <- FindMarkers(allsamples, ident.1 = "HC", ident.2 = "BCR-UV", group.by = 'type', subset.ident = 0)

head(markersingroup)
# Save markers
write.table(markersingroup, file="/data/../TotalSeq/Results/markers/markersingroup.txt", quote=F, row.names = FALSE, sep = "\t")
VlnPlot(allsamples,features=c("HLA-B","IFITM1"), group.by = "type", pt.size = 0)
FeaturePlot(allsamples, features=c("HLA-B"), split.by = "type")
table(allsamples$type)

# Get all the expression count matrix
#head(GetAssayData(allsamples, slot="counts"))
# Get expression values for a single gene from all cells
head(FetchData(allsamples, vars="HLA-B"))

save.image(file=paste(outputpath,"SeuratObjectAfterMarker.RData",sep=""))
sessionInfo()
```

## 7. Generate R script
```{r generatescript, message=FALSE, warning=FALSE}
# Reference
# https://satijalab.org/seurat/articles/multimodal_vignette.html
#file.copy(from = "/data/../TotalSeq/Tools/clustering.html",
#          to   = "/data/../TotalSeq/Reports/site/clustering.html")
#file.remove("/data/../TotalSeq/Tools/clustering.html")
```
